---
title: "DAR F23 Project Status Notebook"
author: "Liebin Zhang"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
subtitle: "Hockey Analytics"
---


## Weekly Work Summary	

**NOTE:** Follow an outline format; use bullets to express individual points. 

* RCS ID: zhangl38
* Project Name: Hockey Analytics
* Summary of work since last week 

    * I dropped the feature of goal and then I build k-means elbow and visualizes means. After this I build Heatmap and visualize the pca graph without goal.

* NEW: Summary of github issues added and worked 

    * Issues that you've submitted
    * Issues that you've self-assigned and addressed
    
* Summary of github commits 

    * include branch name(s)
    * include browsable links to all external files on github
    * Include links to shared Shiny apps
    
* List of presentations,  papers, or other outputs

    * Include browsable links
    
* List of references (if necessary) 
* Indicate any use of group shared code base
* Indicate which parts of your described work were done by you or as part of joint efforts

* **Required:** Provide illustrating figures and/or tables

## Personal Contribution	

* Clearly defined, unique contribution(s) done by you: code, ideas, writing...
* Include github issues you've addressed

Besides the function codes are contained in the file `AnalysisCodeFunc.R`. in my data preparation step that imports as a helper script that defines various functions used through this notebook for data processing and analysis, all other ideas, codes, and writings are all done myself. 

## Analysis: Question 1 

### How to build k-means elbow and visualizes means and build Heatmap without goal?

### Data Preparation


```{r setup_1, echo = FALSE, eval=FALSE}
# RUN THIS CODE CHUNK BEFORE KNITTING (the first time)
# NOTE: Do not run this code when knitting notebook

# Install required packages
r = getOption("repos")
r["CRAN"] = "http://cran.rstudio.com"
options(repos = r)

if (!require("jpeg")) {
   install.packages("jpeg")
}
if (!require("grid")) {
   install.packages("grid")
}
if (!require("scales")) {
   install.packages("scales")
}
if (!require("reshape2")) {
   install.packages("reshape2")
}
if (!require("tidyverse")) {
   install.packages("tidyverse")
}
if (!require("tidymodels")) {
   install.packages("tidymodels")
}
if (!require("ggnewscale")) {
   install.packages("ggnewscale")
}
if (!require("glmnet")) {
   install.packages("glmnet")
}
if (!require("MLmetrics")) {
   install.packages("MLmetrics")
}
if (!require("knitr")) {
   install.packages("knitr")
}
if (!require("knitr")) {
   install.packages("knitr")
}
if (!require("magrittr")) {
   install.packages("magrittr")
}
if (!require("randomForest")) {
   install.packages("randomForest")
}
```


```{r setup_2, include = FALSE}
# Load required R packages
# NOTE: Run this chunk every time but don't include code or results in rendered notebook

# Plotting
library(jpeg)
library(grid)
library(ggnewscale)
library(scales)
# Goal shot stats
library(reshape2)
#library(randomForest)
# library(glmnet)
#library(MASS)
#library(Boruta)
#library(MLmetrics)
#library(kernelshap)
#library(shapviz)

library(knitr)
library(tidyverse)
library(tidymodels)
library(magrittr)
```


#### Setup: Define functions

The major functions of this notebook are contained in the file `AnalysisCodeFunc.R`. This code chunk *sources* (imports) a helper script that defines various functions used through this notebook for data processing and analysis.

```{r setup_3, echo = TRUE}
# All user-defined functions are contained in the following helper script file. 
source("../../AnalysisCodeFunc.R")
```

#### Setting program parameters

This section sets the dimensions of the data structures used in this notebook, based on the captured video.

```{r, echo = TRUE}
# Size of rink image and of all plots
xsize <- 2000
ysize <- 850

# FPS of the video
fps <- 29.97

# Coordinates to the goal pipes
pipes_x <- 1890
lpipe_y <- 395
rpipe_y <- 455
```

#### Data preparation for predictive modelling

Based on our settings, this section reads in the captured image data.

The code is highly dependent upon the following directory structure:

-   The file path determined by the `filepath` variable contains folders named with the game number, followed by 'p', followed by the period number
-   Each `period` folder contains a folder named `Sequences`.
-   Each `Sequences` folder contains `sequence_folders` that contain all the relevant sequence data.

```{r, results='hide', message=FALSE}
# Set filepaths and read the rink
# Results suppressed because this can be messy...

# This file path should contain the hockey rink images and all the sequences
filepath <- '../../FinalGoalShots/'

# See above for explanation of file path syntax
games <- c(24, 27, 33, 34)
# Only take the first and third periods. These are when the opposing team shoots on our goal. Our shots only accounted for about a fifth of shots, and removing them makes the data more consistent
periods <- map(games, ~ str_c(., 'p', c(1, 3))) %>% unlist

# Get the 'Sequences' folder for every period
period_folders <- map(periods, ~ {
  str_c(filepath, ., '/Sequences')
})

# Get every folder inside each 'Sequences' folder
sequence_folders <- period_folders %>%
  map(~ str_c(., '/', list.files(.))) %>%
  unlist 

# Read the rink images and format them to a raster used for graphing
rink_raster <- makeRaster(filepath, 'Rink_Template.jpeg')
half_rink_raster <- makeRaster(filepath, 'Half_Rink_Template.jpeg')

# As every folder is run through the `combinePasses` function, the info.csv file in each sequence folder is read and its contents inserted as a row in info
info <- matrix(0, nrow = 0, ncol = 4) %>% 
  data.frame %>% 
  set_names(c('possessionFrame', 'shotFrame', 'outcome', 'rightHanded'))

# Read in all the sequences
# NOTE: This step takes a long time (minutes)
sequences = sequence_folders %>% map(combinePasses)

# Change outcomes to more verbose names
info$outcome %<>% fct_recode(Goal = 'G', Save = 'GB', 'Defender Block' = 'DB', Miss = 'M')
```

#### Shot statistics retrieval

This section constructs the dataframe used to predict if shots are successful.

```{r}
# Get stats for the shot in every sequence
shots_stats.df <- seq_along(sequences) %>% 
  map_dfr(goalShotStats) %>% 
  # Some models can't use logical data
  mutate_if(is.logical, as.factor)
```

We first combine the shots data with the outcomes vector;

```{r}
# Split data into training and validation sets
outcomes.goal <- (info$outcome == 'Goal') %>% as.numeric %>% as.factor

# Append to shots_stats.df
shots_stats_goal.df <- cbind(shots_stats.df, outcomes.goal)

# Save this dataframe on the file system in case we want to simply load it later (to save time) 
saveRDS(shots_stats_goal.df, "shots_stats_goal.df.Rds")
```

Now let's do some basic classification analysis on the `shots_stats_goal.df` dataset!

We'll use `tidymodels`, part of the **tidyverse**, to split the data into an 80% train/20% test split.

```{r}
#Create training set
set.seed(100)

# Type ?initial_split , ?training , or ?testing in the R console to see how these work!
hockey_split <- initial_split(shots_stats_goal.df, prop = 0.8)
hockeyTrain <- training(hockey_split)
hockeyTest <- testing(hockey_split)

# Check how many observations for each split we have
nrow(hockeyTrain)
nrow(hockeyTest)

# How many features are there
ncol(hockeyTrain)
```


### Analysis: Methods and results


#### Cluster the unsuccessful shots
Step 1: I use new matrix named shotsNum to take off the goals(outcomes) and change the True/False feature to factor
Step 2: Using shotsNum to cluster the unsuccessful shots as HockeyTrain.m and prepare to build k-means elobow

```{r,warning=FALSE}
shots <- readRDS("shots_stats_goal.df.Rds")#Need to make sure this is grabb
library(data.table)
library(mltools)
shotsNum <- shots
shotsNum$goalieScreened = as.factor(shotsNum$goalieScreened)
shotsNum$oppDefenders = as.factor(shotsNum$oppDefenders)
shotsNum$sameDefenders = as.factor(shotsNum$sameDefenders)
shotsNum <- shotsNum[,1:11]
shotsNum <- one_hot(dt = as.data.table(shotsNum))

h.df <- shotsNum %>%
mutate_all(as.numeric)%>% mutate_all(scale) 
hockeyTrain.m <- as.matrix(h.df)

my.pca<-prcomp(hockeyTrain.m,retx=TRUE)
plot(my.pca, type="line")
```

#### Build k-means elbow

Step 1: using wssplot and cluster set(HockeyTrain.m) to build k-means by cluster
Step2: Analysis k-means elbow which is 10
```{r,warning=FALSE}
wssplot <- function(data, nc=25, seed=20){
wss <- data.frame(cluster=1:nc, quality=c(0))
for (i in 1:nc){
set.seed(seed)
wss[i,2] <- kmeans(data, centers=i)$tot.withinss}
ggplot(data=wss,aes(x=cluster,y=quality)) +
geom_line() +
ggtitle("Quality of k-means by Cluster")
}
wssplot(hockeyTrain.m, nc=25)

```

#### Visualizes means and build Heatmap without goal

Step 1: Visualizes means in 7 clusters which clusters of sizes are 4, 15, 15, 3, 32, 14, 22
Step 2: Using ggplot to visualize my cluster means.
Step 3: Analysis Heatmap: The high relavent in heatmap is pcudistance in cluster 2 and cluster 5, puckAngle has high respective in cluster 1 and cluster 1 and 3.
Result: I plan to analysis puck distance influence in Hockey game in the future analysis. 

```{r,warning=FALSE}
library("gplots")
library("ggbiplot")

km <- kmeans(hockeyTrain.m,7)
km

par(mar = c(1, 1, 1, 1))
heatmap.2(km$centers,
scale = "column",
dendrogram = "column",
Colv=FALSE,
cexCol=1.0,
main = "Kmeans Cluster Centers", trace ="none")
```


### Discussion of results

Based on the analysis that I performed, according to the quality of k mean by cluster graph, we can easily see that the elbow was found at 7. So I choose to divide the data set into 7 different clusters, and then visualize them in the form of heatmap. By taking a look into the heatmap, I got the conclusion that the high relavent in heatmap is pcudistance in cluster 2 and cluster 5, puckAngle has high respective in cluster 1 and cluster 1 and 3. 


## Analysis: Question 2 

### How to visualize clusters with PCA without goal?


### Data Preparation

I use the same data set as Analysis Question 1.

### Analysis: Methods and Results  

#### Visualize clusters with PCA without goal

```{r}

# Create a biplot for PC1 and PC2 colored by cluster
t<-1.2*max(abs(my.pca$x[,1:2]))

p <- ggbiplot(my.pca,
choices=c(1,2),
alpha=.1,
scale = 0,
groups=as.factor(km$cluster))
p <- p + scale_colour_hue()
ggtitle('Biplot of PC1 and PC2')+
xlim(-t,t) + ylim(-t,t)
p
```

### Discussion of results

According to the pca graph, we can see that there are a lot of arrows points toward the point sets. For example, "goalie_screen_FALSE" and "oppDefender_1" are pointing into the group 3 meaning that these two features explain the most behaviors in that group. And also, the features of "puckedAngle", "rightHanded", and "goalieAngle" are pointing into the group 6 meaning that these features explain the most behaviors inside group 6. 



## Summary and next steps

For the next step, I am going to use the given function in analysis code to determine the distance of each player and how they reach the goal. 


