---
title: "Exploring and Analyzing Hockey 'Rush' Plays"
subtitle: "DAR Assignment 2 (Fall 2023)"
author: "Your Name Here"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
r = getOption("repos")
r["CRAN"] = "http://cran.rstudio.com"
options(repos = r)

# Load all required packages

library(knitr)
library(ggplot2)
library(dplyr)
library(sp)
library(jpeg)
library(grid)
library(ggnewscale)
library(alkahest)
```

Read in all data and set parameters: 
```{r}
# Hockey rink image
rink_image = readJPEG("FinalGoalShots/Rink_Template.jpeg")
# Format the image to a raster for graphing
rink_raster = rasterGrob(rink_image, width = unit(1,"npc"), height = unit(1,"npc"),interpolate = FALSE)

# Load in all the rush play files
num_rush_plays = 10
#rush_plays = as.list(zeros(num_rush_plays, 1))
rush_plays = as.list(rep(0, num_rush_plays))

rush_indices = 1:num_rush_plays
# Rush play 5 does not exist, so don't try to load it
rush_indices = rush_indices[rush_indices != 5]

# Rush play files are marked as RP(# of play)
prefix = "RP"

for (index in rush_indices){
  # Load rush plays as csv file names "(prefix)(# of play)" from the "Hockey" folder
  # from the home directory
  rush_plays[[index]] = data.frame(read.csv(paste("FinalGoalShots/",prefix,index,".csv", sep=""), header=TRUE, row.names=NULL))
}

# Size of rink image and of all plots
xsize = 2000
ysize = 850
```

The `plotPlayerPaths()` function gives a graph of the paths of the hockey players on the rink, given a rush play data frame. It also removes the `(2000, 0)` points which are used as placeholders when the players are off-frame and interpolates between known frames

The *use_speed_dots* option (`FALSE` by default), when set to `TRUE`, places dots along the paths whose colors represent the velocity of that player.

The *do_smoothing* option (`TRUE` by default), when set to `TRUE`, uses triangle smoothing to smooth the player paths

The plot title, colors of the paths and dots, and which players to use are controllable parameters with default values.

```{r}
# Function that returns distance between two points
getDistance = function(point_1, point_2){
  return(sqrt(sum((point_1-point_2)^2)))
}

# Function used to get a player's xy coordinates along with the frames they happened on
getXYframes = function(rush_play, player_index){
    # Get the current player's x and y coordinate and velocity
    player_xy = rush_play[(2 * player_index):(2 * player_index + 1)]
    
    # Frame counts start at 1 and increment by 10 every row
    frames = 1+10*0:(nrow(player_xy)-1)
    
    return(cbind(player_xy, frames))
}
# Function used to calculate speed in speedFromXY, in feet per second
calculateSpeed = function(index, xy_frames, fps = 29.97){
  # Store dx, dy, and dt in a single row
  delta = xy_frames[index,] - xy_frames[index-1,]
  
  # 1 pixels = 1/10 foot
  dxfeet = delta[1]/10
  dyfeet = delta[2]/10
  
  # 1 frame = 1/fps seconds
  dtseconds = delta[3]/fps
  
  # Pythagorean theorem for distance
  ddistance = sqrt(dxfeet^2+dyfeet^2)
  
  speed = ddistance/dtseconds
}

# Function that takes in a set of x and y coordinates along with frame counts and returns speed for each row
speedFromXY = function(xy_frames, fps = 29.97){
  # xy_frames: A data frame with columns x, y, and frames
  append(0, unlist(lapply(2:nrow(xy_frames), calculateSpeed, xy_frames, fps), use.names = FALSE))
}
# Function that takes in a data frame and smooths the given columns over a time series using triangular smoothing
smoothDataFrameCols = function(data_frame, columns = 1:ncol(data_frame), span = 3){
  data_frame[columns] = lapply(data_frame[columns], function(dfc) {
    smooth_triangular(1:length(dfc), dfc, span)$y
  })
  
  return(data_frame)
}

# Function that returns player paths, excluding the (2000, 0) coordinates from a rush play
getFilteredPaths = function(rush_play, player_indices = 1:10, do_smoothing = FALSE, smooth_span = 3){
    # Filter the paths for placeholder values
  filtered_paths = as.list(zeros(max(player_indices), 1))
  
  for (player_index in player_indices){
    # Get the player's xy coordinates and frames they happened
    player_xyframes = getXYframes(rush_play, player_index)
    
    # Remove the (2000, 0) coordinates and store the results
    filtered_xy = player_xyframes %>% filter(as.vector(player_xyframes[1] < 2000) & 
                                                            as.vector(player_xyframes[2] > 0))
    # Smooth if applicable
    if(do_smoothing){
      filtered_xy = smoothDataFrameCols(filtered_xy, 1:2, smooth_span)
    }
    
    # Remove frame counts, then add speed
    filtered_paths[[player_index]] = filtered_xy %>%
      mutate('sp.{player_index}' := speedFromXY(filtered_xy)) %>%
      select(-frames)
  }
  return(filtered_paths)
}

# Helper function to create geom_paths from a set of xy coordinates
filteredPathToGeomPath = function(player_index, filtered_paths) {
  fp = data.frame(filtered_paths[player_index])
  # Do some weird stuff with get(names) because it doesn't work otherwise
  return(geom_path(data = fp, aes(x = get(names(fp)[1]), y = get(names(fp)[2]), color = paste("Player", player_index))))
}

# Helper function to create geom points from xy coordinates and velocity
filteredPathToGeomPoint = function(player_index, filtered_paths){
  fp = data.frame(filtered_paths[player_index])
  
  # Don't make a dot for the start, where speed = 0
  fp = fp[2:nrow(fp),]
  
  return(geom_point(data = fp, aes(x = get(names(fp)[1]), y = get(names(fp)[2]), color = get(names(fp)[3]))))
}

plotPlayerPaths = function(rush_play, plot_title="Untitled Rush Play", player_indices = 1:10, path_colors = c('black', 'green', 'blue', 'red', 'orange', 'yellow', 'purple', 'gray', 'cyan', 'brown'), use_speed_dots = FALSE, dot_scale = c("#0000FF", "#00FF00"), do_smoothing = FALSE, smooth_span = 3) {
  plot = ggplot() +
    # Superimpose the image of the rink
    annotation_custom(rink_raster, xmin=0, xmax=xsize, ymin=0, ymax=ysize) +
    # Set coordinates
    ylim(0,ysize) +
    xlim(0,xsize) +
    # Set colors for the paths
    scale_color_manual(values=path_colors) +
    # And add labels and the title
    labs(x = "x", y = "y", color = "Player") +
    ggtitle(plot_title) +
    # Center the title
    theme(plot.title = element_text(hjust = 0.5))

  # Grab the filtered paths
  filtered_paths = getFilteredPaths(rush_play, player_indices, do_smoothing, smooth_span)
  
  # Add the filtered paths to the plot as a list
  plot = plot + lapply(player_indices, filteredPathToGeomPath, filtered_paths)
  
  # Add dots marking speed
  if (use_speed_dots){
    # Allow another color scale for the dots
    plot = plot + new_scale_color() + scale_colour_gradient(name = "f/s",low = dot_scale[1], high = dot_scale[2],limits=c(0,40))
    
    # Add all the dots
    plot = plot + lapply(player_indices, filteredPathToGeomPoint, filtered_paths)
  }
  
  return(plot)
}
```

The "pairToPair" function gets the pair-to-pair separation of all players in a rush play as a list of list of data frames. Each data frame has two columns, one for the distance and one for the frames those distances happened. To get the pair-to-pair separation of two specific players, index the overall list with both player's indices: for example, pairToPair(rush_plays[[2]])[[1]][[2]] for the pair-to-pair separation of players 1 and 2.

```{r}
# Function that gets the distance between two players at a certain row
distanceAtRow = function(row, xy_frames_p1, xy_frames_p2){
  return(getDistance(xy_frames_p1[row, 1:2], xy_frames_p2[row, 1:2]))
}
# Function that creates a pair to pair separation for two players
singlePairToPair = function(xy_frames_p1, xy_frames_p2, do_smoothing = FALSE){
  # Abbreviate
  xyf1 = xy_frames_p1
  xyf2 = xy_frames_p2
  
  # Get all indices with (2000, 0) coordinates in either player
  invalid_indices = xyf1[,1] == 2000 | xyf1[,2] == 0 | xyf2[,1] == 2000 | xyf2[,2] == 0
  
  # Exclude each player's invalid indices
  xyf1 = xyf1[!invalid_indices, ]
  xyf2 = xyf2[!invalid_indices, ]
  
  # Smooth if applicable
  if(do_smoothing){
    xyf1 = smoothDataFrameCols(xyf1, columns = c(1, 2))
    xyf2 = smoothDataFrameCols(xyf2, columns = c(1, 2))
  }
  
  # Get the distances at each row into a vector
  distance = sapply(1:nrow(xyf1), distanceAtRow, xyf1, xyf2)
  
  # Keep the frame counts
  result = cbind(distance, xyf1[3])
  
  return(result)
}

pairToPair = function(rush_play, do_smoothing = FALSE){
  # Get the xy coordinates and frames for all players
  # Then, find all pair to pair distance combinations
  lapply(1:10, function(x) { lapply(lapply(1:10, function(y, rush) { getXYframes(rush, y) }, rush_play), singlePairToPair, getXYframes(rush_play, x), do_smoothing) } )
}

# Takes in a separation data frame and converts it to a geom line
separationDFtoGeomLine = function(index, p2p_list, player_pairs){
  return(geom_line(data = p2p_list[[index]], aes(x = frames, y = distance, color = paste(player_pairs[[index]][1], "vs", player_pairs[[index]][2]))))
}

# Takes in a p2p object and graphs the given combinations (... means you can input unlimited arguments, should be vectors with length two representing the player indices you want to compare)
graphPairToPair = function(p2p_obj, ..., line_colors = c('black', 'green', 'blue', 'red', 'orange', 'yellow', 'purple', 'gray', 'cyan', 'brown'), plot_title = "Untitled P2P Separation Graph"){
  # Convert the unlimited arguments to a form we can use
  player_pairs = list(...)
  
  # Turn the two-item vectors into the appropriateseparation data frames
  separation_list = lapply(player_pairs, function(x) {
    return(p2p_obj[[x[1]]][[x[2]]])
  })
  
  plot = ggplot() +
    # Set the colors
    scale_color_manual(values = line_colors) +
    # Label the legend
    labs(color = "Players") +
    # Set and center the title
    ggtitle(plot_title) +
    theme(plot.title = element_text(hjust = 0.5)) +
    # Set x and y bounds to the min and max values
    xlim(0, 301) + 
    ylim(0, getDistance(c(0, 0), c(2000, 850)))
  
  # Create the list of geom_lines
  separation_lines = lapply(seq_along(separation_list), separationDFtoGeomLine, separation_list, player_pairs)
  
  # Add them to the plot
  plot = plot + separation_lines
  
  return(plot)
}
```

```{r}
plotPlayerPaths(rush_plays[[2]], plot_title = "Unsmoothed")
plotPlayerPaths(rush_plays[[2]], player_indices = c(2, 8), plot_title = "Selecting Players")
plotPlayerPaths(rush_plays[[2]], use_speed_dots = TRUE, plot_title = "With Speed Dots")
plotPlayerPaths(rush_plays[[2]], do_smoothing = TRUE, plot_title = "Smoothed")
```
```{r}
p2p_rp2 = pairToPair(rush_plays[[2]])
p2p_rp2_smooth = pairToPair(rush_plays[[2]], do_smoothing = TRUE)

graphPairToPair(p2p_rp2, c(1, 2), c(1, 4), plot_title = "Player 1 vs 2 and 4")
graphPairToPair(p2p_rp2_smooth, c(1, 2), c(1, 4), plot_title = "Player 1 vs 2 and 4, Smoothed")
```

```{r}

```

