---
title: "DAR F23 Hockey Analytics"
author: "Jeff Jung"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
subtitle: "Hockey Analytics"
editor_options: 
  chunk_output_type: inline
---

## Weekly Work Summary

**NOTE:** Follow an outline format; use bullets to express individual points.

-   RCS ID: jungj6

-   Project Name: Hockey Analytics

Summary of work since last week

-   Last week, I have successfully categorized continuous variables into categories, and this week, I did further analysis with categorical data

-   Created bar plots to show the modes categorized variable with saves data

-   Created heat map, which required converting all categorical data into binary variables

-   Heat map effectively showed the correlations between different categories of each feature

NEW: Summary of github issues added and worked

N/A

Summary of github commits

-   One commit of HW4 to dar-jungj6

List of presentations, papers, or other outputs

N/A

-   List of references (if necessary)

-   Indicate any use of group shared code base

-   Indicate which parts of your described work were done by you or as part of joint efforts

-   **Required:** Provide illustrating figures and/or tables

## Personal Contribution

-   Analysis on the modes of failure with categorization
-   Application of bar plots in illustrating different modes of failure with categorization
-   Application of heat map in illustration different modes of failure with categorization

## Analysis: Modes of Failure

### Question being asked

Since we do not have much data for goals, I have decided to do further analysis on saves. Finding the modes of failure would allow me to identify what features of shots led to saves.

### Data Preparation

Data preparation for hockeyTrain is the same as I did for other assignments. Since I am specifically examining the modes of failure, I have separated 'saves' from hockeyTrain. Also, I have categorized continuous variables into 3 groups (high, med, low), so that it is easier to visualize.

```{r, hockeytrain, include=FALSE}
# Include all data processing code (if necessary), clearly commented
# Include all data processing code (if necessary), clearly commented

# Install required packages
r = getOption("repos")
r["CRAN"] = "http://cran.rstudio.com"
options(repos = r)

if (!require("jpeg")) {
   install.packages("jpeg")
}
if (!require("grid")) {
   install.packages("grid")
}
if (!require("scales")) {
   install.packages("scales")
}
if (!require("reshape2")) {
   install.packages("reshape2")
}
if (!require("tidyverse")) {
   install.packages("tidyverse")
}
if (!require("tidymodels")) {
   install.packages("tidymodels")
}
if (!require("ggnewscale")) {
   install.packages("ggnewscale")
}
if (!require("glmnet")) {
   install.packages("glmnet")
}
if (!require("MLmetrics")) {
   install.packages("MLmetrics")
}
if (!require("knitr")) {
   install.packages("knitr")
}
if (!require("knitr")) {
   install.packages("knitr")
}
if (!require("magrittr")) {
   install.packages("magrittr")
}

# Plotting
library(jpeg)
library(grid)
library(ggnewscale)
library(scales)
# Goal shot stats
library(reshape2)
library(knitr)
library(tidyverse)
library(tidymodels)
library(magrittr)

# All user-defined functions are contained in the following helper script file. 
source("../../AnalysisCodeFunc.R")

# Size of rink image and of all plots
xsize <- 2000
ysize <- 850

# FPS of the video
fps <- 29.97

# Coordinates to the goal pipes
pipes_x <- 1890
lpipe_y <- 395
rpipe_y <- 455

# This file path should contain the hockey rink images and all the sequences
filepath <- '../../FinalGoalShots/'

# See above for explanation of file path syntax
games <- c(24, 27, 33, 34)
# Only take the first and third periods. These are when the opposing team shoots on our goal. Our shots only accounted for about a fifth of shots, and removing them makes the data more consistent
periods <- map(games, ~ str_c(., 'p', c(1, 3))) %>% unlist

# Get the 'Sequences' folder for every period
period_folders <- map(periods, ~ {
  str_c(filepath, ., '/Sequences')
})

# Get every folder inside each 'Sequences' folder
sequence_folders <- period_folders %>%
  map(~ str_c(., '/', list.files(.))) %>%
  unlist 

# Read the rink images and format them to a raster used for graphing
rink_raster <- makeRaster(filepath, 'Rink_Template.jpeg')
half_rink_raster <- makeRaster(filepath, 'Half_Rink_Template.jpeg')

# As every folder is run through the `combinePasses` function, the info.csv file in each sequence folder is read and its contents inserted as a row in info
info <- matrix(0, nrow = 0, ncol = 4) %>% 
  data.frame %>% 
  set_names(c('possessionFrame', 'shotFrame', 'outcome', 'rightHanded'))

# Read in all the sequences
# NOTE: This step takes a long time (minutes)
sequences = sequence_folders %>% map(combinePasses)

# Change outcomes to more verbose names
info$outcome %<>% fct_recode(Goal = 'G', Save = 'GB', 'Defender Block' = 'DB', Miss = 'M')

# Get stats for the shot in every sequence
shots_stats.df <- seq_along(sequences) %>% 
  map_dfr(goalShotStats) %>% 
  # Some models can't use logical data
  mutate_if(is.logical, as.factor)

# Split data into training and validation sets
outcomes.goal <- (info$outcome == 'Goal') %>% as.numeric %>% as.factor

# Append to shots_stats.df
shots_stats_goal.df <- cbind(shots_stats.df, outcomes.goal)

# Save this dataframe on the file system in case we want to simply load it later (to save time) 
saveRDS(shots_stats_goal.df, "shots_stats_goal.df.Rds")

#Create training set
set.seed(100)

# Type ?initial_split , ?training , or ?testing in the R console to see how these work!
hockey_split <- initial_split(shots_stats_goal.df, prop = 0.8)
hockeyTrain <- training(hockey_split)
hockeyTest <- testing(hockey_split)

# Check how many observations for each split we have
nrow(hockeyTrain)
nrow(hockeyTest)

# How many features are there
ncol(hockeyTrain)
```

```{r, result01_data}
# Load necessary library
library(dplyr)

# Calculate quantiles to divide the data into thirds
puckDist_q <- quantile(hockeyTrain$puckDist, probs = c(1/3, 2/3))
puckAngle_q <- quantile(hockeyTrain$puckAngle, probs = c(1/3, 2/3))
puckSpeed_q <- quantile(hockeyTrain$puckSpeed, probs = c(1/3, 2/3))
shooterSpeed_q <- quantile(hockeyTrain$shooterSpeed, probs = c(1/3, 2/3))
goalieDist_q <- quantile(hockeyTrain$goalieDist, probs = c(1/3, 2/3))
goalieAngle_q <- quantile(hockeyTrain$goalieAngle, probs = c(1/3, 2/3))
posTime_q <- quantile(hockeyTrain$posTime, probs = c(1/3, 2/3))

# Create three categories based on quantiles
hockeyTrain <- hockeyTrain %>%
  mutate(puckSpeedCategory = case_when(
    puckSpeed <= puckSpeed_q[1] ~ "Low",
    puckSpeed <= puckSpeed_q[2] ~ "Medium",
    TRUE ~ "High"
  ))

hockeyTrain <- hockeyTrain %>%
  mutate(puckDistCategory = case_when(
    puckDist <= puckDist_q[1] ~ "Low",
    puckDist <= puckDist_q[2] ~ "Medium",
    TRUE ~ "High"
  ))

hockeyTrain <- hockeyTrain %>%
  mutate(puckAngleCategory = case_when(
    puckAngle <= puckAngle_q[1] ~ "Low",
    puckAngle <= puckAngle_q[2] ~ "Medium",
    TRUE ~ "High"
  ))

hockeyTrain <- hockeyTrain %>%
  mutate(shooterSpeedCategory = case_when(
    shooterSpeed <= shooterSpeed_q[1] ~ "Low",
    shooterSpeed <= shooterSpeed_q[2] ~ "Medium",
    TRUE ~ "High"
  ))

hockeyTrain <- hockeyTrain %>%
  mutate(goalieDistCategory = case_when(
    goalieDist <= goalieDist_q[1] ~ "Low",
    goalieDist <= goalieDist_q[2] ~ "Medium",
    TRUE ~ "High"
  ))

hockeyTrain <- hockeyTrain %>%
  mutate(goalieAngleCategory = case_when(
    goalieAngle <= goalieAngle_q[1] ~ "Low",
    goalieAngle <= goalieAngle_q[2] ~ "Medium",
    TRUE ~ "High"
  ))

hockeyTrain <- hockeyTrain %>%
  mutate(posTimeCategory = case_when(
    posTime <= posTime_q[1] ~ "Low",
    posTime <= posTime_q[2] ~ "Medium",
    TRUE ~ "High"
  ))

# Subset the data for not goals (e.g., "Save")
saves_data <- hockeyTrain[hockeyTrain$outcomes.goal != "1", ]
```


### Analysis: Methods and results

First, I have created a bar chart to show what the mode groups are for each variable. And then, to show how each feature is correlated with other features, I have created a heatmap.

```{r, result01_analysis}
# Include all analysis code, clearly commented
# If not possible, screen shots are acceptable. 
# If your contributions included things that are not done in an R-notebook, 
#   (e.g. researching, writing, and coding in Python), you still need to do 
#   this status notebook in R.  Describe what you did here and put any products 
#   that you created in github. If you are writing online documents (e.g. overleaf 
#   or google docs), you can include links to the documents in this notebook 
#   instead of actual text.

# Find modes for each variable
mode_puckSpeed <- names(sort(table(saves_data$puckSpeedCategory), decreasing = TRUE)[1])
mode_puckDist <- names(sort(table(saves_data$puckDistCategory), decreasing = TRUE)[1])
mode_puckAngle <- names(sort(table(saves_data$puckAngleCategory), decreasing = TRUE)[1])
mode_shooterSpeed <- names(sort(table(saves_data$shooterSpeedCategory), decreasing = TRUE)[1])
mode_goalieDist <- names(sort(table(saves_data$goalieDistCategory), decreasing = TRUE)[1])
mode_goalieAngle <- names(sort(table(saves_data$goalieAngleCategory), decreasing = TRUE)[1])
mode_posTime <- names(sort(table(saves_data$posTimeCategory), decreasing = TRUE)[1])

# Print the modes for each variable
cat("Mode for puckSpeedCategory:", mode_puckSpeed, "\n")
cat("Mode for puckDistCategory:", mode_puckDist, "\n")
cat("Mode for puckAngleCategory:", mode_puckAngle, "\n")
cat("Mode for shooterSpeedCategory:", mode_shooterSpeed, "\n")
cat("Mode for goalieDistCategory:", mode_goalieDist, "\n")
cat("Mode for goalieAngleCategory:", mode_goalieAngle, "\n")
cat("Mode for posTimeCategory:", mode_posTime, "\n")

# Define custom colors for each category
category_colors <- c("Low" = "lightgreen", "Medium" = "lightyellow", "High" = "lightcoral")

# Create a bar plot
ggplot(saves_data, aes(x = puckSpeedCategory, fill = puckSpeedCategory)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = category_colors) +  # Set custom colors
  labs(
    title = "Puck Speed Category by Outcome",
    x = "Puck Speed Category",
    y = "Count",
    fill = "Puck Speed Category"  # Legend title
  )
ggplot(saves_data, aes(x = puckDistCategory, fill = puckDistCategory)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = category_colors) +  # Set custom colors
  labs(
    title = "Puck Dist Category by Outcome",
    x = "Puck Dist Category",
    y = "Count",
    fill = "Puck Dist Category"  # Legend title
  )
ggplot(saves_data, aes(x = puckAngleCategory, fill = puckAngleCategory)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = category_colors) +  # Set custom colors
  labs(
    title = "Puck Angle Category by Outcome",
    x = "Puck Angle Category",
    y = "Count",
    fill = "Puck Angle Category"  # Legend title
  )
ggplot(saves_data, aes(x = shooterSpeedCategory, fill = shooterSpeedCategory)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = category_colors) +  # Set custom colors
  labs(
    title = "Shooter Speed Category by Outcome",
    x = "Shooter Speed Category",
    y = "Count",
    fill = "Shooter Speed Category"  # Legend title
  )
ggplot(saves_data, aes(x = goalieDistCategory, fill = goalieDistCategory)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = category_colors) +  # Set custom colors
  labs(
    title = "Goalie Dist Category by Outcome",
    x = "Goalie Dist Category",
    y = "Count",
    fill = "Goalie Dist Category"  # Legend title
  )
ggplot(saves_data, aes(x = goalieAngleCategory, fill = goalieAngleCategory)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = category_colors) +  # Set custom colors
  labs(
    title = "Goalie Angle Category by Outcome",
    x = "Goalie Angle Category",
    y = "Count",
    fill = "Goalie Angle Category"  # Legend title
  )
ggplot(saves_data, aes(x = posTimeCategory, fill = posTimeCategory)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = category_colors) +  # Set custom colors
  labs(
    title = "Pos Time Category by Outcome",
    x = "Pos Time Category",
    y = "Count",
    fill = "Pos Time Category"  # Legend title
  )
```

### Discussion of results

The results are pretty reasonable. I see that there is a tie between low and medium for puck angle, and some of other variables have a small difference between two categories. But this is due to lack of data, and it looks like this method can explain a lot about modes of failure.

## Analysis: Heat Map of the Modes of Failures

### Question being asked

We have analyzed the modes of failure. Now, I want to find out the relativity between one mode of failure with other modes of failures. More specifically, I want to find out what features are common and correlated when a shot has failed. Also, we will further examine by categorizing continous variables.

### Data Preparation

First, I have created a subset of data which takes in continous variables. Then I found correlation between those varibles and converted it into a matrix so that it can be used for making the heatmap. To make an another heatmap for categorical data, since the heatmap can only take numerical values, I have first converted categorical variables that were continous into binary variables and made a subset of these variables. Then I found the correlation of these variables and graphed into a heatmap.

```{r, result02_data}
# Include all data processing code (if necessary), clearly commented

# Define the features that need to be included
keeps <- c("puckDist", "puckAngle", "puckSpeed", "shooterSpeed", "goalieDist", "goalieAngle")

# Create a data frame for the heatmap with 
heatmap_data <- saves_data[keeps]

# Find correlation
heatmap_matrix <- cor(heatmap_data[sapply(heatmap_data, is.numeric)])

# Converts categorical variables with 3 categories to binary binary variables
saves_data <- saves_data %>%
  mutate(
    puckSpeedHigh = ifelse(puckSpeedCategory == "High", 1, 0),
    puckSpeedMedium = ifelse(puckSpeedCategory == "Medium", 1, 0),
    puckSpeedLow = ifelse(puckSpeedCategory == "Low", 1, 0),
    
    puckDistHigh = ifelse(puckDistCategory == "High", 1, 0),
    puckDistMedium = ifelse(puckDistCategory == "Medium", 1, 0),
    puckDistLow = ifelse(puckDistCategory == "Low", 1, 0),
    
    puckAngleHigh = ifelse(puckAngleCategory == "High", 1, 0),
    puckAngleMedium = ifelse(puckAngleCategory == "Medium", 1, 0),
    puckAngleLow = ifelse(puckAngleCategory == "Low", 1, 0),
    
    shooterSpeedHigh = ifelse(shooterSpeedCategory == "High", 1, 0),
    shooterSpeedMedium = ifelse(shooterSpeedCategory == "Medium", 1, 0),
    shooterSpeedLow = ifelse(shooterSpeedCategory == "Low", 1, 0),
    
    goalieDistHigh = ifelse(goalieDistCategory == "High", 1, 0),
    goalieDistMedium = ifelse(goalieDistCategory == "Medium", 1, 0),
    goalieDistLow = ifelse(goalieDistCategory == "Low", 1, 0),
        
    goalieAngleHigh = ifelse(goalieAngleCategory == "High", 1, 0),
    goalieAngleMedium = ifelse(goalieAngleCategory == "Medium", 1, 0),
    goalieAngleLow = ifelse(goalieAngleCategory == "Low", 1, 0)
  )

# Subset of data to be represented in heatmap
catkeeps <- c("puckSpeedHigh", "puckSpeedMedium", "puckSpeedLow", "puckDistHigh", "puckDistMedium", "puckDistLow", "puckAngleHigh", "puckAngleMedium", "puckAngleLow", "shooterSpeedHigh", "shooterSpeedMedium", "shooterSpeedLow", "goalieDistHigh", "goalieDistMedium", "goalieDistLow", "goalieAngleHigh", "goalieAngleMedium", "goalieAngleLow")

# Now we have added binary variables to saves_data, we can take subset of saves_data
heatmap_catdata <- saves_data[catkeeps]

# Find correlation between binary categorical variables
heatmap_catmatrix <- cor(heatmap_catdata[sapply(heatmap_catdata, is.numeric)])
```

### Analysis: Methods and Results

All data have been prepared above, and heatmap_matrix and heatmap_catmatrix are in appropriate matrix form only with numeric values. We can use the heatmap function to produce a heatmap.

```{r, result02_analysis}
# Include all analysis code, clearly commented
# If not possible, screen shots are acceptable. 
# If your contributions included things that are not done in an R-notebook, 
#   (e.g. researching, writing, and coding in Python), you still need to do 
#   this status notebook in R.  Describe what you did here and put any products 
#   that you created in github. If you are writing online documents (e.g. overleaf 
#   or google docs), you can include links to the documents in this notebook 
#   instead of actual text.

# Heatmap without categorization
heatmap(heatmap_matrix)

# Heatmap with the categorization
heatmap(heatmap_catmatrix)
```

### Discussion of results

We can clearly see that some variables are highly correlated, and there exist variables that are not so correlated to other variables. What is more interesting is that the correlation between the same variables in different categories can vary a lot. For example, we can see that the correlation between puckAngleHigh and goalieAngleHigh is very high, meaning that a shot within these two categories is likely to be a save. However, the correlation between puckAngleHigh and goalieAngleLow is very low, meaning a shot within these two categories is likely to not be a save.

## Analysis: N/A

### Question being asked

### Data Preparation

```{r, result03_data}
# Include all data processing code (if necessary), clearly commented

```

### Analysis methods used

*Describe in natural language a statement of the analysis you're trying to do*

*Provide clearly commented analysis code; include code for tables and figures!*

```{r, result03_analysis}
# Include all analysis code, clearly commented
# If not possible, screen shots are acceptable. 
# If your contributions included things that are not done in an R-notebook, 
#   (e.g. researching, writing, and coding in Python), you still need to do 
#   this status notebook in R.  Describe what you did here and put any products 
#   that you created in github. If you are writing online documents (e.g. overleaf 
#   or google docs), you can include links to the documents in this notebook 
#   instead of actual text.
```

### Discussion of results

*Provide in natural language a clear discussion of your observations.*

## Summary and next steps

Next week, I will find some way to incorporate categorical data into cluster analysis.
