---
title: "DAR F23 Project Status Notebook Template"
author: "Student Name"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
subtitle: "DAR Project Name Here ('LLM Fairness' or 'DeFi LTM' or 'Hockey Analytics'"
---

## Instructions (DELETE BEFORE SUBMISSION)

* Use this notebook is a template for your biweekly project status assignment. 
* Use the sections starting with **BiWeekly Work Summary** as your outline for your submitted notebook.
* Summarize ALL of your work in this notebook; **if you don't show and/or link to your work here, it doesn't exist for us!**

1. Create a new copy of this notebook in the `AssignmentX` sub-directory of your team's github repository using the following naming convention

   * `rcsid_assignmentX.Rmd` and `rcsid_assignmentX.pdf`
   * For example, `bennek_assignment03.Rmd`

2. Document **all** the work you did on your assigned project this week **using the outline below.** 

3. You MUST include figures and/or tables to illustrate your work. *Screen shots are okay*, but include something!

4. You MUST include links to other important resources (knitted HTMl files, Shiny apps). See the guide below for help.

5. Commit the source (`.Rmd`) and knitted (`.html`) versions of your notebook and push to github

6. **Submit a pull request.** Please notify Dr. Erickson if you don't see your notebook merged within one day. 

7. **DO NOT MERGE YOUR PULL REQUESTS YOURSELF!!**

See the Grading Rubric for guidance on how the contents of this notebook will be graded on LMS or GradeScope. 

## Weekly Work Summary	

**NOTE:** Follow an outline format; use bullets to express individual points. 

* RCS ID: **Always** include this!
* Project Name: **Always** include this!
* Summary of work since last week 

    * Describe the important aspects of what you worked on and accomplished

* NEW: Summary of github issues added and worked 

    * Issues that you've submitted
    * Issues that you've self-assigned and addressed
    
* Summary of github commits 

    * include branch name(s)
    * include browsable links to all external files on github
    * Include links to shared Shiny apps
    
* List of presentations,  papers, or other outputs

    * Include browsable links
    
* List of references (if necessary) 
* Indicate any use of group shared code base
* Indicate which parts of your described work were done by you or as part of joint efforts

* **Required:** Provide illustrating figures and/or tables

## Personal Contribution	

* Clearly defined, unique contribution(s) done by you: code, ideas, writing...
* Include github issues you've addressed

## Analysis: Question 1 (Provide short name)

### Question being asked 

_Provide in natural language a statement of what question you're trying to answer_

### Data Preparation

_Provide in natural language a description of the data you are using for this analysis_

_Include a step-by-step description of how you prepare your data for analysis_

_If you're re-using dataframes prepared in another section, simply re-state what data you're using_

```{r, result01_data}
# Include all data processing code (if necessary), clearly commented

```

### Analysis: Methods and results

_Describe in natural language a statement of the analysis you're trying to do_

_Provide clearly commented analysis code; include code for tables and figures!_

```{r, result01_analysis}
# Include all analysis code, clearly commented
# If not possible, screen shots are acceptable. 
# If your contributions included things that are not done in an R-notebook, 
#   (e.g. researching, writing, and coding in Python), you still need to do 
#   this status notebook in R.  Describe what you did here and put any products 
#   that you created in github. If you are writing online documents (e.g. overleaf 
#   or google docs), you can include links to the documents in this notebook 
#   instead of actual text.

```

### Discussion of results

_Provide in natural language a clear discussion of your observations._


## Analysis: Question 2 (Provide short name)

### Question being asked 

_Provide in natural language a statement of what question you're trying to answer_

### Data Preparation

_Provide in natural language a description of the data you are using for this analysis_

_Include a step-by-step description of how you prepare your data for analysis_

_If you're re-using dataframes prepared in another section, simply re-state what data you're using_

```{r, result02_data}
# Include all data processing code (if necessary), clearly commented

```

### Analysis: Methods and Results  

_Describe in natural language a statement of the analysis you're trying to do_

_Provide clearly commented analysis code; include code for tables and figures!_

```{r, result02_analysis}
# Include all analysis code, clearly commented
# If not possible, screen shots are acceptable. 
# If your contributions included things that are not done in an R-notebook, 
#   (e.g. researching, writing, and coding in Python), you still need to do 
#   this status notebook in R.  Describe what you did here and put any products 
#   that you created in github. If you are writing online documents (e.g. overleaf 
#   or google docs), you can include links to the documents in this notebook 
#   instead of actual text.

```

### Discussion of results

_Provide in natural language a clear discussion of your observations._


## Analysis: Question 3 (Provide short name)

### Question being asked 

_Provide in natural language a statement of what question you're trying to answer_

### Data Preparation

_Provide in natural language a description of the data you are using for this analysis_

_Include a step-by-step description of how you prepare your data for analysis_

_If you're re-using dataframes prepared in another section, simply re-state what data you're using_

```{r, result03_data}
# Include all data processing code (if necessary), clearly commented

```

### Analysis methods used  

_Describe in natural language a statement of the analysis you're trying to do_

_Provide clearly commented analysis code; include code for tables and figures!_

```{r, result03_analysis}
# Include all analysis code, clearly commented
# If not possible, screen shots are acceptable. 
# If your contributions included things that are not done in an R-notebook, 
#   (e.g. researching, writing, and coding in Python), you still need to do 
#   this status notebook in R.  Describe what you did here and put any products 
#   that you created in github. If you are writing online documents (e.g. overleaf 
#   or google docs), you can include links to the documents in this notebook 
#   instead of actual text.

```


### Discussion of results

_Provide in natural language a clear discussion of your observations._

## Summary and next steps

_Provide in natural language a clear summary and your proposed next steps._

